---
title: "Leontief Inverse Calculation"
output: 
  html_document:
    toc: true
    toc_float: true
---

# Bangladesh Leontief Inverse 

Load required packages 
```{r}
library(data.table)
library(readxl)
```

## Data Cleaning 

#### Read in data 

Let's read in the Bangladesh IO table
```{r}
setwd("/Users/prodyumna_goutam/Downloads")
input_file <- read_excel("Inpur Outpur Table _2007_Final.xls",
                      sheet="IOT07", 
                      skip = 1)
bgd_iot <- as.data.table(input_file)
```

#### Clean up

Take a quick look at what the io table looks like 
```{r}
head(bgd_iot)
```

We need to do a few data cleaning things:
```{r}
bgd_iot <- bgd_iot[-1,] # drop first obs thats just names of sectors
colnames(bgd_iot) <- paste0("io",colnames(bgd_iot)) # prepend colnames with IO 

dropcol <- c("io..89","io..90","io..91",
             "io..92","io..93","io..94",
             "io..95","io..96")
bgd_iot[, (dropcol) := NULL] # drop unnecessary columns 
```

Let's take a look at the last few rows bgd_iot table  
```{r}
bgd_iot[86:96,1:5]
```

Notice that it contains the Value Added summary measure that we will need for our calculation, so extract that and drop the summary measures because we won't need it 
```{r}
bgd_iot_summ <- bgd_iot[`ioMillion Taka` %in% c("Value Added"), ] # extract value added 
bgd_iot_summ[, c("io..1","ioMillion Taka") := NULL]

bgd_iot <- bgd_iot[!(87:101), ] # drop unnecessary summary measures 
bgd_iot[,`ioMillion Taka` := NULL] # drop unnecessary rows
setnames(bgd_iot,"io..1","Commodity") # rename columns 
bgd_iot[, Commodity := paste0("io",Commodity)]
```


Convert columns to numeric from character
```{r}
bgd_iot[,c(paste0("io",1:86))] <- lapply(bgd_iot[,c(paste0("io",1:86))], as.numeric)
bgd_iot_summ <- bgd_iot_summ[, lapply(.SD,as.numeric), .SDcols = paste0("io",1:86)]
```


Finally, convert to a matrix 
```{r}
bgd_iot_mat <- as.matrix(bgd_iot, rownames = "Commodity")
bgd_iot_summ_mat <- as.matrix(bgd_iot_summ)
```


## IO Coefficient matrix 

We now have an io table with input usage, we will need to convert this to an IO coeffcient matrix. 

This is fairly straightforward: 

1. Add across rows of the io table to get total input usage 
2. Add the value added for each industry 
3. Divide each cell of input table by the total output for a particular sector 


Let's write a function to do that 

```{r}
io_coeff_calc <- function(iomat,suppmat) {
  
  #Sum across rows for total input
  totinput <- colSums(iomat)  
  
  # Create total output
  totoutput <- vector(mode = "numeric", length = length(totinput))
  names(totoutput) <- names(totinput)
  for (i in names(totoutput)){
     totoutput[i] <- totinput[i] + suppmat[,i]
  }
  
  # IO Coefficient matrix 
  # Note: We are going to be explicit and index by the 
  # the name of the sector and loop
  iot_coeff <- matrix(nrow = nrow(iomat), ncol = ncol(iomat))
  rownames(iot_coeff) <- rownames(iomat)
  colnames(iot_coeff) <- colnames(iomat)

  for (i in colnames(iot_coeff)) {
    iot_coeff[,i] <- iomat[,i]/totoutput[i]
  }
  return (iot_coeff)
}
```


```{r}
bgd_iot_coeff <- io_coeff_calc(bgd_iot_mat,bgd_iot_summ_mat)
```


## Leontief inverse

```{r}
# Identity matrix 
ident <- diag(nrow = nrow(bgd_iot_coeff))

# Leontief Inverse
bgd_li <- solve(ident - bgd_iot_coeff)
```


Read out as csv 
```{r}
write.csv(bgd_li, file = "bgd_li.csv")
```


# India Leontief Inverse 

## Data Cleaning 

#### Read in data
```{r}
input_file <- read_excel("India IO Matrix 2003-04.xls",
                         skip = 5)
ind_iot <- as.data.table(input_file)
```

#### Clean up

Take a look 
```{r}
head(ind_iot)
```


```{r}
tail(ind_iot)
```


As above, let's do a little cleanup 
```{r}
# Drop unnecessary columns 
dropcol <- c("IIUSE","PFCE","GFCE",
             "GFCF","CIS","EXP",
             "less IMP","TFUSE","Total")
ind_iot[, (dropcol) := NULL]
setnames(ind_iot,c("Commodity / Industry","..2"),c("IndCode","Industry"))
```


We also want to separate out the summary rows at the bottom of the data and use it separately 
```{r}
rlist <- c("Total Input","NIT","Total Input+NIT", 
           "GVA","Total Output")
ind_iot_summ <- ind_iot[Industry %in% rlist, ][, IndCode := NULL]
ind_iot <- ind_iot[!(Industry %in% rlist), ]
```



## India-Bangladesh IO Concordance 

So, it's clear that the organization of the India IO is different that the organization of the Bangladesh IO that we constructed in the section above. 

We will need to build a concordance and sum the rows and columns of the India IOT appropriately. 


First, load in the concordance between India and Bangladesh IO sectors
```{r}
input_file <- read_excel("Exports and Informality Mappings Final.xlsx", 
                         sheet = "INDIO-BGDIO")
ind_bgd <- as.data.table(input_file)
```

We will now need to sum the rows. Let's write a function to do that. 
```{r}
sum_dt <- function(x,y,by_x,by_y) {
  merge_dt <- merge(x,y,by.x=by_x,by.y=by_y,all=TRUE) # Merge dt with mapping 
  sum_merge_dt <- merge_dt[ , lapply(.SD,sum), by = consolid_iot_code]
  return(sum_merge_dt)
}
```

Sum rows of the india iot
```{r}
ind_iot_merge <- sum_dt(ind_iot[, !c("Industry")], 
                  ind_bgd[,c("ind_iot_code","consolid_iot_code")],
                  by_x = "IndCode",
                  by_y = "ind_iot_code") 
```


Now, we will sum the columns using a small trick: we will transpose the data and then sum the rows of the transposed table as above 

Transpose first 
```{r}
ind_iot_mergeT <- dcast(melt(ind_iot_merge,id.vars = "consolid_iot_code", variable.name = "IndCode"), 
                        IndCode ~ consolid_iot_code)
ind_iot_mergeT[, IndCode := as.numeric(as.character(IndCode))]
```

OKay, let's repeat what we did before 

```{r}
ind_iot_mergeT <- merge(ind_iot_mergeT,
                       ind_bgd[,!c("ind_iot_name","consolid_iot_name")], 
                       by.x = "IndCode",by.y = "ind_iot_code", all = TRUE)
ind_iot_mergeT <- ind_iot_mergeT[, lapply(.SD,sum), by = consolid_iot_code]
ind_iot_mergeT[, IndCode := NULL]
```


Rotate Back to get the final consolidated Indian IO table 

```{r}
ind_iot_consolid <- dcast(melt(ind_iot_mergeT,id.vars = "consolid_iot_code"), 
                        variable ~ consolid_iot_code)
```


Okay, let's convert this to a matrix and proceed 

```{r}
ind_iot_mat <- as.matrix(ind_iot_consolid, rownames = "variable")
```



## IO Coefficient Matrix 

Let's take a look at the summary matrix which contains the India information: 

```{r}
head(ind_iot_summ)
```

We need the GVA (Gross Value Added) row, but notice that the column codes are India IOT codes


Notice that the columns actually follow the coding for India IOT. We need to consolidate this as well:



```{r}
ind_iot_summT <- dcast(melt(ind_iot_summ, 
                            id.vars = "Industry", 
                            variable.name = "IndCode",
                            variable.factor = FALSE),
                       IndCode ~ Industry)
```


```{r}
ind_iot_summT[,IndCode := as.numeric(IndCode)]
setnames(ind_iot_summT,"GVA","ValueAdded")
ind_iot_summT <- merge(ind_iot_summT[,c("IndCode","ValueAdded")],
                       ind_bgd[,c("ind_iot_code","consolid_iot_code")],
                       by.x = "IndCode",by.y = "ind_iot_code",
                       all = TRUE)
ind_iot_summT <- ind_iot_summT[, lapply(.SD,sum), by = consolid_iot_code]
ind_iot_summT[, consolid_iot_code := as.character(consolid_iot_code)]
```


```{r}
totinput <- colSums(ind_iot_mat)
```

```{r}
extractcols <- names(totinput)
```

```{r}
totoutput <- vector(mode = "numeric", length = length(totinput))
names(totoutput) <- extractcols
for (i in extractcols){
   totoutput[i] <- totinput[i] + ind_iot_summT[consolid_iot_code==i,ValueAdded]
}
```

```{r}
iot_coeff <- matrix(nrow = nrow(ind_iot_mat), ncol = ncol(ind_iot_mat))
rownames(iot_coeff) <- rownames(ind_iot_mat)
colnames(iot_coeff) <- colnames(ind_iot_mat)

for (i in colnames(iot_coeff)) {
  iot_coeff[,i] <- ind_iot_mat[,i]/totoutput[i]
}
```


```{r}
ident <- diag(nrow = nrow(iot_coeff))
ind_li <- solve(ident - iot_coeff)
```



```{r}
write.csv(ind_li, file = "ind_li.csv")
write.csv(iot_coeff, file = "ind_iot_coeff.csv")
```

